---
title: "Homework 1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
by: Emily 

## Problem 1 CH2 ISLR p.9
```{r}
#reading in data
AutoData=read.csv('/Users/admin/Documents/Csvv/Auto.csv',header=TRUE)

```

```{r}
#looking at the type of data we have for each variable
lapply(AutoData, class)
```

The data set has Horsepower characterized as a char because of the "?", so we convert it to numeric and omit the NAs.   

```{r}
#Converting into Numeric to get NAs and subsetting No Na data 
AutoData$horsepower=as.numeric(AutoData$horsepower)
AutoNew=na.omit(AutoData)
```
AutoNew is the modified data set we will be using. 

<br>

**(a)**  
```{r}
#attaching new data
attach(AutoNew)
lapply(AutoNew,class)
```
Quantitative predictors include mpg,displacement,horsepower, weight, year, and origin. Qualitative includes the name of the Auto, and cylinders. 

<br>

**(b)**  
```{r}
# Finding range for each variable
range(mpg)
range(displacement)
range(horsepower)
range(weight)
range(acceleration)
range(year)
range(origin)
```

<br>

**(c)** 
```{r, echo=FALSE}
#finding mean and standard deviation for each variable

#mpg
m1=mean(mpg)
s1=sd(mpg)

#displacement
m2=mean(displacement)
s2=sd(displacement)

#horsepower
m3=mean(horsepower)
s3=sd(horsepower)

#weight
m4=mean(weight)
s4=sd(weight)

#acceleration
m5=mean(acceleration)
s5=sd(acceleration)

#year
m6=mean(year)
s6=sd(year)

#origin
m7=mean(origin)
s7=sd(origin)
```

```{r,echo=FALSE}
#Adding in statements to explain what the mean and sd is for each variable

cat('The mean for mpg is' ,m1, ' and the sd is',s1,'\n')
cat('The mean for displacement is' ,m2, ' and the sd is',s2,'\n')
cat('The mean for horsepower is' ,m3, ' and the sd is',s3,'\n')
cat('The mean for weight is' ,m4, ' and the sd is',s4,'\n')
cat('The mean for acceleration is' ,m5, ' and the sd is',s5,'\n')
cat('The mean for year is' ,m6, ' and the sd is',s6,'\n')
cat('The mean for origin is' ,m7, ' and the sd is',s7,'\n')
```

<br>

**(d)**

```{r}
#removing rows as wanted from textbook
RemoveRowsData=AutoNew[-c(10:84),]
```

```{r,echo=FALSE}
#creating function to display the mean, sd, range for each variable in new subset
Temp=function(var1,var2){
  m=mean(var1)
  sd1=sd(var1)
  ranger=range(var1)
  srt=as.character(var2)
  cat('The mean for',srt,'is' ,m, ', the sd is',sd1, 'and the range is',ranger,'\n')
}
Temp(RemoveRowsData$mpg,"mpg")
Temp(RemoveRowsData$displacement,"displacement")
Temp(RemoveRowsData$horsepower,"horsepower")
Temp(RemoveRowsData$weight,"weight")
Temp(RemoveRowsData$acceleration,"acceleration")
Temp(RemoveRowsData$year,"year")
Temp(RemoveRowsData$origin,"origin")

```

<br>

**(e)**
```{r}
#Pairs plot
pairs(~mpg+displacement+horsepower+weight+acceleration+year+origin, data=AutoNew)
```

There is positive linear correlation among some of the predictors such as displacement and horsepower, horsepower and weight, and lastly displacement and weight. 

<br>

**(f)**
There appears to be multiple predictors that seem to have a negative correlation with mpg such as displacement, horsepower and weight. It could be possible to predict mpg with those variables. 


## Problem 2: nonISLR analyzing data

```{r}
#loading in data
Parasite=read.table('/Users/admin/Documents/Csvv/CodParasite.txt',header=TRUE)
```

```{r}
#preforming log transformation on 2nd column 
Parasite$Intensity=Parasite[,2]+1
Parasite$Intensity=log(Parasite[,2])
```


```{r}
#creating function to make boxplots 
BP=function(C1,C2,strC2){
  str=as.character(strC2)
  boxplot(C1~C2,data=Parasite,xlab= str,ylab="Intensity (# of Parasites)",main="Boxplot")
}

BP(Parasite$Intensity,Parasite$Area,"Area")
BP(Parasite$Intensity,Parasite$Sex,"Sex")
BP(Parasite$Intensity,Parasite$Stage,"Stage")
BP(Parasite$Intensity,Parasite$Year,"Year")

```

The box plot of Area shows the most association to intensity compared to the other variables. According to the plot most of the data in the different areas from 1 to 3 are skewed and have possible multiple outliers. Area 4 shows an increase in median, since the data is more spread out. The plots do have overlap but some are longer than others. 

The box plot for years also seems to have a bit of an association with intensity as the year 2000 seems to have have increased, then in 2001 the median fell.


## Problem 3, nonISLR analyzing data

```{r}
#loading in data
Owls=read.table('/Users/admin/Documents/Csvv/Owls.txt',header=TRUE)
```

<br>

**(a)**

```{r}
#Factoring categorical variables 
Owls$SexParent=as.factor(Owls$SexParent)
Owls$FoodTreatment=as.factor(Owls$FoodTreatment)
```


```{r}
#attaching data
attach(Owls)
sapply(Owls,class)
```


```{r}
#creating boxplots
boxplot(NegPerChick~SexParent)
boxplot(NegPerChick~FoodTreatment)
boxplot(NegPerChick~SexParent+FoodTreatment+FoodTreatment*SexParent,data=Owls)
```

<br>

**(b)**
```{r}
boxplot(NegPerChick~Nest)
```

<br>

There are differences for Neg per chick across the various nesting sites. 

**(c)**
```{r}
plot(NegPerChick~SiblingNegotiation)
```

<br>

Yes,there is an association between these features as it seems like negperchick increases as sibling negation increases. 

<br>

**(d)**
```{r}
#Log transformation for graphing purposes 
Owls$SiblingNegotiation=Owls[,5]+1
Owls$SiblingNegotiation=log(Owls[,5])
plot(Owls$SiblingNegotiation~Owls$ArrivalTime)
```




