---
title: "STA 4364 HW4"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Emily 

<br>

## Information: 

Predicting tumor type from gene expression data. We will build a multinomial regression model with l1 and l2. 

\newline

## Part a 

```{r}
#loading in datasets

GeneData=read.csv('/Users/admin/Documents/Csvv/gene_data.csv',header=TRUE)
GeneLabels=read.csv('/Users/admin/Documents/Csvv/gene_labels1.csv',header=TRUE,
                    stringsAsFactors = TRUE)


```

<br>

```{r}
#loading in libraries 
library(glmnet)
library(stats)
library(caret)
library(dplyr)

```

<br>

```{r}
#looking for NAs in coloumns 
sum(is.na(GeneData))

```

<br>

```{r}
#Finding variance from columns and removing variance with less than 0.001
#NumericCol=GeneData[sapply(GeneData,is.numeric)]

VarianceCol=sapply(GeneData,var)
removeVar=which(VarianceCol<0.001)
NewGeneData=GeneData[,-removeVar]

```

<br>


```{r}
# Standardize each gene predictor column to have mean 0, and sd 1.
StdGeneData=
    NewGeneData[2:dim(NewGeneData)[2]]  %>%
 mutate_all(~(scale(.) %>% as.vector))
```

<br>

```{r}
#Add in X Col to merge the datasets 
StdGeneData$X=rep(0,length(StdGeneData$gene_0))
StdGeneData$X=NewGeneData$X

```


<br>

```{r}
#merge the datasets 
GeneDataFull=merge(StdGeneData,GeneLabels, by="X",all=TRUE)
GeneDataFull=GeneDataFull[,-c(1)]
str(GeneDataFull)

```

<br>

```{r}
#splitting data into training and testing
n_train_samps = ceiling(0.75 * nrow(GeneDataFull))
train_inds = sample(1:nrow(GeneDataFull), n_train_samps, replace=FALSE)
val_inds = setdiff(1:nrow(GeneDataFull), train_inds)
```

<br>

```{r}
# train set 
Gene_x_train = GeneDataFull[train_inds, 1:(ncol(GeneDataFull)-1)] # no last col
Gene_x_train_mat =as.matrix(Gene_x_train)

Gene_y_train = GeneDataFull[train_inds,ncol(GeneDataFull)] # last column

```

<br>

```{r}
# test set
Gene_x_val = GeneDataFull[val_inds, 1:(ncol(GeneDataFull)-1)] # no last col
Gene_y_val = GeneDataFull[val_inds,ncol(GeneDataFull)] # last column
```

<br>

##Part b

\newline

Ridge Logistic Regression with 10 fold CV

```{r}
#Creating a ridge logistic regression model with 10 fold CV
gene.ridge=cv.glmnet(x=Gene_x_train_mat, y=Gene_y_train, nfolds = 10,
                     alpha=0,family = "multinomial")
```

<br>

```{r}
#plotting coefficients
plot(gene.ridge$glmnet.fit, label=TRUE)
```

<br>


```{r}
#plotting CV error 
plot(gene.ridge)
```
\newline 

<br>

```{r}
#Finding best lambda value
bestlam=gene.ridge$lambda.min
bestlam
```

<br>

```{r}
#creating prediction model on the feature validation set
ridge.pred = predict(gene.ridge, s =bestlam, newx = as.matrix(Gene_x_val) , type="class")

```

<br>

```{r}
#making confusion matrix from prediction model and testing response set
confusionMatrix(table(ridge.pred,Gene_y_val))
```

<br>

##Part c

\newline

LASSO Logistic Regression with 10 fold CV

```{r}
#creating Lasso model
gene.lasso = cv.glmnet(x=Gene_x_train_mat, y=Gene_y_train, nfolds = 10,
                     alpha=1,family = "multinomial")

```

<br>

```{r}
#Optimal value of lambda
bestlamLasso=gene.lasso$lambda.min
```


```{r}
#Prediction model using feature testing set
lasso.pred=predict(gene.lasso, s =bestlamLasso, newx = as.matrix(Gene_x_val),type="class")

```


```{r}
#confusion matrix
confusionMatrix(table(lasso.pred,Gene_y_val))

```

<br>

## Part d

```{r}

lasso.coef = predict(gene.lasso, as.matrix(Gene_x_val), type ="coefficients", s = bestlamLasso)
```


```{r}
#genes associated with BRCA 
which(coef(gene.lasso)$BRCA>0)
```

<br>


```{r}
#genes associated with COAD
which(coef(gene.lasso)$COAD>0)
```


<br>

```{r}
#genes associated with KIRC
which(coef(gene.lasso)$KIRC>0)
```

<br>

```{r}
#genes associated with LUAD
which(coef(gene.lasso)$LUAD>0)
```

<br>

```{r}
#genes associated with PRAD
which(coef(gene.lasso)$PRAD>0)

```

<br>

**Top 20 Relevant Genes **:

\newline

1. Gene 1857
2. Gene 9150
3. Gene 15835 
4. Gene 18328 
5. Gene 357 
6. Gene  6590
7. Gene 219
8. Gene  9892 
9. Gene 220
10. Gene 11973 
11. Gene 3735
12. Gene 3438 
13. Gene 18455
14. Gene 2827
15. Gene  5560
16. Gene 13770
17. Gene 14042
18. Gene 6714
19. Gene 17844
20. Gene 13251

















